---
title: "Class12: Structural Bioinformatics 1"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Section 1: PDB Statistics
Here we inspect the types of structures in the main databases for 3D bimolecular data - the PDB.

> Q1: Determine the percentage of structures solved by X-Ray and Electron Microscopy. 

```{r}
#read in file
stats <- read.csv("Data Export Summary.csv", row.names = 1)

# Calculate the percentage values for each 'method' asked for
ans <- (stats$Total/sum(stats$Total))*100
names(ans) <- rownames(stats)
round(ans, 2)
```
X-ray: 88.95% and Electron microscopy: 2.72%

> Q2: Also can you determine what proportion of structures are protein? Aim to have a rendered GitHub document with working code that yields your answers.

```{r}
protein <- sum(stats$Proteins)/sum(stats$Total)*100
round(protein, 2)
```
92.69%

> Q3: Type HIV in the PDB website search box on the home page and determine how many HIV-1 protease structures are in the current PDB?

There are 1289 HIV-1 protease structures in the PDB database.

## Section 2: Using VMD
Used the VMD application to create a structure of HIV that is saved as hiv_pic.tiff.tga

## Hands-on Part 2: 

First we have to install all the packages using this code:
"install.packages("bio3d")
install.packages("ggplot2")
install.packages("ggrepel")
install.packages("devtools")
install.packages("BiocManager")

BiocManager::install("msa")
devtools::install_bitbucket("Grantlab/bio3d-view")"

Read a single PDB structure into R
```{r}
library(bio3d)
pdb <- read.pdb("1hsg")
pdb
```

What is this pdb object that is stored? It is a list of 8 things.
```{r}
attributes(pdb) # use this to see the headers
pdb$atom
```

Let's write a new file that just has the protein only:
```{r}
pdb <- read.pdb("1HSG")
ligand <- atom.select(pdb, "ligand", value = TRUE) #this function chooses the ligand only from the pdb file which would be for the drug
protein <- atom.select(pdb, "protein", value = TRUE) #for the protein
```

Now we want to write out these new PDB objects with the `write.pdb()` function:
```{r}
write.pdb(ligand, file = "1hsg_ligand.pdb")
write.pdb(protein, file = "1hsg_protein.pdb")
```


