---
title: "Class12: Structural Bioinformatics 1"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## PDB Statistics
Here we inspect the types of structures in the main databases for 3D bimolecular data - the PDB.

> Q1: Determine the percentage of structures solved by X-Ray and Electron Microscopy. 

```{r}
#read in file
stats <- read.csv("Data Export Summary.csv", row.names = 1)

# Calculate the percentage values for each 'method' asked for
ans <- (stats$Total/sum(stats$Total))*100
names(ans) <- rownames(stats)
round(ans, 2)
```

> Q2: Also can you determine what proportion of structures are protein? Aim to have a rendered GitHub document with working code that yields your answers.

```{r}
protein <- sum(stats$Proteins)/sum(stats$Total)*100
round(protein, 2)
```

> Q3: Type HIV in the PDB website search box on the home page and determine how many HIV-1 protease structures are in the current PDB?

There are 1289 HIV-1 protease structures in the PDB database.
