---
title: "Class 13: Structure based drug discovery"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Download and clean HIV-pr receptor structure
Here we will download the 1HSG PDB structure from the online database, remove the bound drug from Merk and write out a protein only and ligand only new set of PDB files.

We will use the **bio3d package** functions `get.pdb()`, `read.pdb()`, `atom.select()`, and finally `write.pdb()`.
```{r}
library(bio3d)
file.name <- get.pdb("1hsg")
#Download PDB file
hiv <- read.pdb(file.name)
```
You can get a quick summary of the pdb structure by calling hiv:
```{r}
hiv
```
> Q1: What is the name of the two non protein resid values in this structure? What does resid correspond to and how would you get a listing of all reside values in this structure?
  - the names are HOH (127) and MK1 (1)

Now make the two protein and ligand files. Select protein and ligand only:
```{r}
prot <- atom.select(hiv, string = "protein", value = TRUE)
write.pdb(prot, file = "1hsg_protein.pdb")

lig <- atom.select(hiv, string = "ligand", value = TRUE)
write.pdb(lig, file = "1hsg_ligand.pdb")
```

To look at the new files you just made:
```{r}
lig
prot
```




