---
title: 'Class 6: R Functions'
author: "Cindy Tran"
date: "1/23/2020"
output: 
  github_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Function

## My level 2 heading

### My level 3 heading

```{r}
# This is a silly plot!
plot(1:5)
```

# Start of today's class

Lets see more about **file import** (i.e. reading files into R). The main read function iin base R is `read.table()`

Help for `read.table` can be brought up as a ?

Read all of the test files:
Test1
normally:
read.table("test1.txt", header = TRUE, sep = ","
If you do read.csv() you don't need to put in all the extra stuff.
```{r}
t1 <- read.csv("test1.txt")
t1
```


Test2
```{r}
t2 <- read.table("test2.txt", header = TRUE, sep = "$")
t2
```

Test3
```{r}
t3 <- read.table("test3.txt")
t3
```
The default for `read.table()` is `header=FALSE`

# Back to functions

Our first example function:
```{r}
add <- function(x, y=1) {
  # Sum of the input x and y
  x + y
}
```
Lets try using this function:
```{r}
add(7)
add(7, 3)
```
How does this work with vectors:
```{r}
add(c(1, 2, 4))
```
It gives you back a vector!
```{r}
add(c(1,2,4),4)
```
This adds 4 to the vectors because the y value is now 4.
```{r}
add(c(1, 2, 4), c(1, 2, 4))
```
The first vector is your x and the second is your y so it adds the repective locations together. 

Can't do `add(c(1, 2, 4), 4, 5, 6)` because it is more than 2 arguments. It's more than just an x and y.

Moving on.... 
What is this range function?
```{r}
x <- c(4,4,10,3, 11)
max(x)
min(x)
range(x)
```
It gives you both the min and the max of the vector. Min is first and max is second.

Test the function from the slides (Function 2):
```{r}
rescale <- function(x) {
   rng <-range(x)
   (x - rng[1]) / (rng[2] - rng[1])
}
rescale(x)
```
Test it on the value of 1:10:
```{r}
rescale(1:10)
```
How would it work here:
```{r}
rescale(c(1,2,NA,3,10))
```
but you want to get a value, not NA's...
This is why you get NA's:
```{r}
y <- c(1,2,NA,3,10)
rng <-range(y)
rng
```
You can fix it by doing this:
```{r}
y <- c(1,2,NA,3,10)
rng <-range(y, na.rm=TRUE)
rng
```
`na.rm=TRUE` means that you are removing any NA from being considered in your function

So now what...
```{r}
rescale2 <- function(x) {
   rng <-range(x, na.rm = TRUE)
   (x - rng[1]) / (rng[2] - rng[1])
}
rescale2(c(1,2,NA,3,10))
```
Now it works and it skips the NA value!

If you type `rescale` into your console region it will return the description of your code telling you what the function was. 

Lets do some more complicated stuff...
```{r}
 rescale3 <- function(x, na.rm=TRUE, plot=FALSE) {
    rng <-range(x, na.rm=na.rm)
    print("Hello")
   answer <- (x - rng[1]) / (rng[2] - rng[1])
   print("is it me you are looking for?")
   if(plot) {
     print("Don't sing again please!")
      plot(answer, typ="b", lwd=4)
   }
   print("I can see it in ...")
   return(answer)
 }
```
Lets try this:
```{r}
rescale3(x)
```
We added the "dont sing again!" print:
```{r}
rescale3(y, plot=TRUE)
```

# Handout
Section B
First we installed the package in the console section because it only needs to be installed one time:
used `install.packages("bio3d")`

Now we will call it so we can use everything in it:
```{r}
library(bio3d)
```

Paste everything from the handout into here:
```{r}
library(bio3d)
s1 <- read.pdb("4AKE")  # kinase with drug
s2 <- read.pdb("1AKE")  # kinase no drug
s3 <- read.pdb("1E4Y")  # kinase with drug
s1.chainA <- trim.pdb(s1, chain="A", elety="CA")
s2.chainA <- trim.pdb(s2, chain="A", elety="CA")
s3.chainA <- trim.pdb(s3, chain="A", elety="CA")
s1.b <- s1.chainA$atom$b
s2.b <- s2.chainA$atom$b
s3.b <- s3.chainA$atom$b
plotb3(s1.b, sse=s1.chainA, typ="l", ylab="Bfactor")
plotb3(s2.b, sse=s2.chainA, typ="l", ylab="Bfactor")
plotb3(s3.b, sse=s3.chainA, typ="l", ylab="Bfactor")
```
This is complicated. We have to figure out what this code does so that we can shorten this. Lets make a function for this for **HOMEWORK**:

Q1: It pulls up data into the Global Environment section from PDB about the protein you chose. Its a large list of 8 things and of class "pdb", "sse".
Q2:It "produces a new smaller PDB object that contains a subset of atoms from a given larger PDB object". It trims the larger PDB file giving a smaller version of the information on the protein.
Q3: The parameter "sse" in the plot code removes the black and gray marginal rectangles. They represent the secondary structure object returned from dssp, stride or in certain cases read.pdb. In dssp, it could be a helix, sheet, or turn.  
Q4: A better plot is plotting the three plots on one on top of one another. 
```{r}
plotb3(s1.b, sse=s1.chainA, typ="l", ylab="Bfactor")
points(s2.b, typ="l", col="blue", lwd=2)
points(s3.b, typ="l", col= "red", lwd=2)
```
Q5: The s2.b and s3.b are more similar to eachother in their B-factor because they are on the same branch in the cluster dendrogram. "dist" quantifies the distance between the structures so you could use that value to see how similar they are to one another. 
```{r}
hc <- hclust( dist( rbind(s1.b, s2.b, s3.b) ) )
plot(hc)
```

Q6:
```{r}
 library(bio3d) #This line calls the library that the proteins are coming from. 
drug <- function(x) {
  s <- read.pdb(x)
  chainA <- trim.pdb(s, chain="A", elety="CA")
  b <- chainA$atom$b
  plotb3(b, sse=chainA, typ="l", ylab = "Bfactor")
}
drug("4AKE")
drug("1AKE")
drug("1E4Y")
```

